\sys: Controllable Program Regeneration

### 2025-05-22 paper/draft.tex
small fix
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

The following query for \sys uses lines-of-code a coarse metric of idiomaticity.
\sys will regenerate the program, minimizing this metric.
\begin{wrapfigure}[3]{r}{.5\columnwidth}
  \vspace{-10pt}
  \begin{minted}{prolog}
equal(io(P), io(P_)).
#min len(P_).
  \end{minted}
\end{wrapfigure}
\sys extracts the $\{\ttt{sig(F)}, \ttt{io(F)}\}$ properties from the original program, where \texttt{sig(F)} is the
signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
on a set of generated inputs.
In this case, where it is instructed to optimize an aspect of the program, \sys will re-prompt
the model to regenerate the program until no improvements are made or there is a negligible difference after two consecutive iterations.
After \xxx seconds, and \xxx iterations, \sys produces the following program:
\begin{minted}[frame=lines]{c}
#include <math.h>
float Q_rsqrt(float number) {
    return 1.0f / sqrtf(number);
}
\end{minted}
The regenerated version is easier to verify, portable across compilers, and
avoids undefined behavior.
A user could have also provided performance requirements using 
the \ttt{time(P)} property to further guide the regeneration process.
In this case, \sys would have run the two functions inside a tight loop and 
would have aborted the regeneration process if the performance of the generated
program was not within an acceptable user-defined threshold.

@@

These results support the claim that \sys’s planning logic — driven by explicit property extraction and knowledge-base reasoning — yields safer and more controllable transformations than prompt-only LLM-based systems.

>>>>

The following query for \sys uses lines-of-code a coarse metric of
idiomaticity. \sys will regenerate the program, minimizing this metric, while
preserving its I/O behavior.
\begin{wrapfigure}[3]{r}{.5\columnwidth}
  \vspace{-5pt}
  \begin{minted}{prolog}
equal(io(P), io(P_)).
#min len(P_).
  \end{minted}
\end{wrapfigure}
\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{io(F)}\}$ from the original program, where \texttt{sig(F)} is the
signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
on a set of generated inputs.
In this case, where it is instructed to optimize an aspect of the program, \sys will re-prompt
the model to regenerate the program until no improvements are made or there is a negligible difference after two consecutive iterations.
After \xxx seconds, and \xxx iterations, \sys produces the following program:
\begin{minted}[frame=lines]{c}
#include <math.h>
float Q_rsqrt(float number) {
    return 1.0f / sqrtf(number);
}
\end{minted}
The regenerated version is easier to verify, portable across compilers, and
avoids undefined behavior.
A user could have also provided performance requirements using 
the \ttt{time(P)} property to further guide the regeneration process.
In this case, \sys would have run the two functions inside a tight loop and 
would have aborted the regeneration process if the performance of the generated
program was not within an acceptable user-defined threshold.

@@

These results support the claim that \sys’s planning logic---driven by explicit
property extraction and knowledge-base reasoning---yields safer and more
controllable transformations than prompt-only LLM-based systems.
